### 2015-05-12 reports/paper/experiment.tex
paper update

\section{Experiment}
To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid. Fig~\ref{fig:setup_point} shows a picture of the setup.

@@

A total of $32$ positions are chosen uniformly in this region where microphone data is recorded. To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error lowers as window size increases and plataeus after window size exceeds around $10$ milli-second. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used to estimate difference of delay

Although accuracy improves with window length, the calculation time also increases with window length. The part of calculation that depends on window length is using cross correlation to estimate difference of arrival time. cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$. We measured how the computation time varies with window length and Figure~\ref{fig:speed_vs_window} shows the result. The runtime increases approximately linearly in the window size region of interest.

@@

Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate arrival time differences. The localization is noisy. As window size increases, the localization converges to the shape of ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.

@@

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was chosen

\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was used.

@@

Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does depend on how fast the sound sorce is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B. Localization accuracy is the best for white noise source, and worst for Music B. This is consistent with our expectation because low amplitude regions in Music B would make arrays lose track of where the source is. This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}

Fig ~\ref{fig:circle_normal} also shows that raw detection has the most amount of jiggling. Kalman filter reduces the amount of jiggling from raw detection. Averaging filter has the least amount of jiggling. However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lags the real movement.

>>>>

\section{Experiment Results}
To test localization accuracy, an one meter by one meter grid was set up and the arrays are placed at the top left and top right corners of the grid. Fig~\ref{fig:setup_point} shows a picture of the setup.

@@

A total of $32$ positions are chosen uniformly in this region where microphone data is recorded. To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error lowers as window size increases and plataeus after window size exceeds around $10$ milli-second. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.

Although accuracy improves with window length, the calculation time also increases with window length. The part of calculation that depends on window length is using cross correlation to estimate TDOA. cross correlation can be calculated with FFT and the runtime is of order $O(N\log N)$. We measured how the computation time varies with window length and Figure~\ref{fig:speed_vs_window} shows the result. The runtime increases approximately linearly in the window size region of interest.

@@

Fig~\ref{fig:err_win_circle} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA. The localization is noisy. As window size increases, the localization converges to the shape of ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. The error decreases as window size increases and plateaus after window size exceeds around $10$ milliseconds.

@@

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.

\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.

@@

Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B. Localization accuracy is the best for white noise source, and worst for Music B. This is consistent with our expectation because low amplitude regions in Music B would make arrays lose track of where the source is. This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}.

Fig ~\ref{fig:circle_normal} also shows that raw detection has the most amount of jiggling. Kalman filter reduces the amount of jiggling from raw detection. Averaging filter has the least amount of jiggling. However, averaging filter averages detection outputs from past 0.5 seconds, which makes the filtered output lag the real movement.

### 2015-05-12 reports/paper/system.tex
paper update

\section{System}
The system uses two arrays, each with three microphones. Three microphones along with a micro-controller are mounted at three vertices of an equilateral triangle. Fig~\ref{fig:setup_array} shows a picture of the array. Micro-controller used in this project is \emph{teensy 3.1}. The micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{array_close.JPG}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{array.JPG}
  \end{subfigure}
  \caption{array}
  \label{fig:setup_array}
\end{figure}

From a timing point of view, the micro controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing. Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.

>>>>

\section{System}
The system uses two arrays, each with three microphones mounted on vetices of a $20$cm equilateral triangle. A micro-controller is also attached to one of the vertices. Fig~\ref{fig:setup_array} shows a picture of the array setup. Micro-controller used in this project is \emph{teensy 3.1}. It has $64$k RAM memory and the ADC is capable of sampling at $500$kHz. In this project, the micro-controller collects microphone data on all three channels for $12$ millisecond and then send the recorded data to a computer for localization.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{array_close.JPG}
    \caption{micro-controller}
  \end{subfigure}
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{array.JPG}
    \caption{array}
  \end{subfigure}
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{setup_2.JPG}
    \caption{two arrays}
  \end{subfigure}
  \caption{Localization arrays}
  \label{fig:setup_array}
\end{figure}

To speed up computation for real time localization, instead of searching for $t_0$ that maximizes equation~\ref{eq:gcc}, a grid search in 2D grid is performed. For each point in the grid, the TDOA to each microphone pair can be precomputed. Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid. To further improve localization accuracy, instead of using point estimate for TDOA, a liklihoood map is built. Each entry in GCC output is used as the liklihood for that delay. With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$. TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map can be built as:
\begin{eqnarray*}
  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_2}(\tau)$, and $R_{m_1,m_2}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Liklihood maps from two arrays can be combined into final liklihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represents the liklihood map from array $1$ and array $2$.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{left.png}
    \caption{localization with only array 1}
    \label{fig:liklihood1}
  \end{subfigure}
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{right.png}
    \caption{localization with only array 2}
    \label{fig:liklihood2}
  \end{subfigure}
  \begin{subfigure}[]{.15\textwidth}
    \includegraphics[width=\textwidth]{combined.png}
    \caption{localization with both arrays}
    \label{fig:liklihood3}
  \end{subfigure}
  \caption{Liklihood maps for localizating point at $(0.0,-0.3)$ m}
  \label{fig:liklihood}
\end{figure}

To see the effect of using multiple arrays, fig~\ref{fig:liklihood} shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual array gives accurate angle estimate, but has high uncertainty in distance estimate. By combining estimates from two arrays, the angle estimate can be effectively combined to estimate distance.

From a timing point of view, the micro-controller spends $12$ millisecond on sampling microphone data before sending the data to the computer for processing. Sending data through USB port takes another $18$ millisecond, and processing on computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.
